### 2021-06-22 README.Rmd
First public version of data package

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# idealista18

<!-- badges: start -->
<!-- badges: end -->

The package idealista18 includes real estate data for three major markets in Spain (Barcelona, Madrid, and Valencia) for the year 2018.

## Installation

<!--
You can install the released version of idealista18 from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("idealista18")
```
-->

You can install the package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paezha/idealista18")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(idealista18)
## basic example code
```