### 2022-02-05 data-in-brief/Tables and Figures.Rmd

---
title: "Tables and Figures"
author: "Fernando & Compay"
date: "2/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Descriptive table

```{r total, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
H <-rbind(H,colSums(H))
text_tbl <- data.frame(
City = c("Barcelona","Madrid","Valencia","Total Ads"),
H,
check.names = FALSE)

xtable::xtable(text_tbl, digits=0, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some dummy variables has been excluded of this table for save space. \\label{tab:Dummy-variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "12em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

```{r table-dummy-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(4,5,6,7,27,30,31,37,38,39)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("Asking price per m^2 (euros)","Surface (m^2)",
                       "Number of bedrooms","Number of bathrooms",
                       "Año construcción","Construction year",
                       "Max build floor",
                       "Distance to city center","Distance to subway station","Distance to main street"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables])),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables])),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some dummy variables has been excluded of this table for save space. \\label{tab:Dummy-variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "12em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

```{r table-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(8,9,10,12,13,15)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("=1 if has terrace","=1 if has lift","=1 if has AA",
                       "=1 if has parking","=1 if has parking include in price","=1 if has north orientation"),
# Barcelona
mean = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Madrid
mean = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Valencia
"mean"=colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of dummy varaibles. See the help facility in the \\textbf{idealista18} package for details and formal definitions \\label{tab:Variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "18em", italic = FALSE) %>%
column_spec(2, width = "16em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```