### 2022-07-12 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Copied decision to `editorial-materials` in preparation to begin revising the paper

---
title: "Using machine learning to identify spatial market segments: A reproducible study of major Spanish markets - Response to referees"
bibliography: bibliography.bib
output: pdf_document
---

<!--
```{r install-data-package, include = FALSE}
# Run only once if needed to install data package `sobiEquity`
if (!require("idealista18", character.only = TRUE)) {
      devtools::install_github("https://github.com/paezha/idealista18")
  }
```

```{r setup,  cache=FALSE, warning = FALSE, include=FALSE, message = FALSE}
library(idealista18)
library(tidyverse)

set.seed(1234)
```
-->

> Dear Authors, Thank you for the opportunity to review your paper. This manuscript is about the publishing of a one-year relocated and randomised listing price data set for three cities in Spain. Both the asking price and coordinates information are anonymised which limits the usefulness of the data for teaching and academic research. Given the house price data is partly from the Idealista company, it is not clear under which kind of licence the data is being issued and to what extent it may be considered open. The unknown quality of this one year randomised data is the main issue for potential users. Having read the paper many times, I believe that the authors have created the idealista18 package (https://github.com/paezha/idealista18). The Idealista18 Github readme file is too simple to guidance users on how to access the data with the R package. So I am not sure it is the right one. The authors are to be commended for their intention to make open real estate data available in an R package which should render it easily accessible for teaching and research. Below is a list of comments and questions that I had which may need to be addressed in future editions of this work.

> 1.	Please reference the package and dataset and include in the text and in the package the license under which the data is released. The Idealista18 github site is unclear on how to access the data with the package.

> 2.	Please state your anticipated research impacts from the publication of this data set in the introduction.  It would be useful to state how this one-year relocated, and randomised house price data might be best used by academics in research and teaching. The relocated and randomised data may be difficult to further enrich with other variables, the authors need to show how the 44 variables is enough for future users.

>3.	The method of randomisation is described but it would be advantageous to publish the code for this so that researchers can evaluate the impact of this kind of randomisation on this and other data sets. Does the method, as seems likely, have a differential effect on data from high density areas as compared with low?
To what extent are analyses likely to be confined to neighbourhood levels of resolution given the method of spatial randomisation? It would be useful to have some idea of the size of the neighbourhoods to facilitate comparison with other spatial data sets

>4.	More details on the data collection methods of Idealista would be useful in interpreting the data. Does Idealista concentrate on particular market segments or not? Are the list prices good indicators of actual sale prices? Are there sources from which actual paid prices could be identified?

>5.	It would be useful for the authors to evaluate the datasets in the light of the FAIR guiding principles for research data stewardship (findability, accessibility, interoperability, and reusability).

>6.	Given the asking price has resized the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences quantitative research results and findings with this data.

>7.	Please kindly add in a usage notes section how this data can be useful for the EPB audience. For example, introduce how this anonymised house price data can be integrated with other data and research.
The original data could be a nice data set for a series of housing related papers. If the authors are unable to publish this original asking price, it maybe worthwhile to publish the research based on this data rather than using the randomised data. House price is locationally sensitive, but both the location and price data in this manuscript are altered by anonymisation. The authors need to show for which analyses the data remains valid despite anonymisation.

# References