### 2022-07-29 anonymization/record_anonymization.Rmd
Draft of comment to reviewer

---
title: 'Listings anonymization (price and coordinates)'
output: pdf_document
---

# Introduction

This markdown file contains the code used for the anonymization of the package's listing records (due to legal requirement). It introduces a small random disturbance in the total price and exact position.

# Set parameters

\tiny
```{r Load libraries, echo=F}
knitr::opts_chunk$set(eval = F)
knitr::opts_chunk$set(size="small")
```
\normalsize

Set process parameters

\tiny
```{r Set process parameters}
dir_geom_path <- paste0(here::here(),"/data/wkt/")  
dir_path <- paste0(here::here(),"/data/assets/raw/")
output_path <- str_remove(dir_path,  "raw/")

#
# Process parameters (expected from a PARAM list)
# here we define some settings
#
COUNTRY = 'es'     # Spain 
TYPOLOGY = 'home'  # This type refers to flats/apartments
OPERATION = 'rent' # This refers to rental
YEAR = 2018
OUTPUT_FORMAT = 'csv.gz'   # Format used for the generated data set
CLEAN_CITYNAME = 'Madrid' # City name: Madrid, Barcelona or Valencio

knitr::opts_chunk$set(echo = TRUE)
```
\normalsize

We set the used CRS, input parameters are coded using and EPSG:4326 and we need to use a projected CRS, in this case we will use "+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs".
\tiny
```{r Set process CRS}
#
# Set the projections used
#
projcrs_src = '+proj=longlat +datum=WGS84 +no_defs'
projcrs_dest = '+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs'
```
\normalsize

# Load data

Load input data (with actual coordinates and prices).
\tiny
```{r Load polygons, message=F, warning=F}
#
# Set input file names
#

file_name = paste0(CLEAN_CITYNAME,'_polygons.',OUTPUT_FORMAT)

#
# Load file and project to make the spatial transformations
#

polygons.sf <- read_delim(paste0(dir_geom_path, file_name), 
                          delim = ";") |>
  st_as_sf(wkt = 'WKT', 
           crs = projcrs_src) |>
  st_transform(projcrs_dest)

#
# Plot geometry
#

plot(st_geometry(polygons.sf))
```
\normalsize

# Load and preprocess listing data

## Load data

Load data and remove some extreme values (taken from idealista history data).
\tiny
```{r Data load, message=F, warning=F}
#
# Variable boundaries come are defined as valid values 
#
max_unitprice = list('sale' = 10443.64, 'rent' = 47.3)
min_price = list('sale' = 5000, 'rent' = 100)
max_price = list('sale' = 4E6, 'rent' = 20E3)
range_area = c(20, 500)

file_name = paste0(COUNTRY,'_',TYPOLOGY,
                   '_',OPERATION,'_',CLEAN_CITYNAME,
                   '_',YEAR,'.',OUTPUT_FORMAT)

# Load individual records
assets.sf <- read_delim(paste0(dir_path, file_name),  delim = ";")

```
\normalsize

## Fix cadastral construction year

Given we take as right the year user fills in we need to fix it, if we see extreme values < 1500 or in the future we assign the median year in the area:
\tiny
```{r Fix construction year, message=F, warning=F}
#
# Impute construction year if it is out of bounds ]-inf, MIN_YEAR] or [YEAR, inf]
#

MIN_YEAR = 1500
imputed_years.sf <- assets.sf |> 
  dplyr::group_by(LOCATIONID, .drop=F) |>
  dplyr::filter(CADCONSTRUCTIONYEAR > MIN_YEAR & CADCONSTRUCTIONYEAR <= YEAR) |>
  dplyr::summarise(IMPUTED_YEAR=as.integer(median(CADCONSTRUCTIONYEAR), .groups=NULL))

assets.sf <- assets.sf |>
  dplyr::inner_join(imputed_years.sf, by='LOCATIONID') |>
  # 1st attempt: Impute with idealista construction year
  dplyr::mutate(CADCONSTRUCTIONYEAR = ifelse(CADCONSTRUCTIONYEAR < MIN_YEAR 
                                          | CADCONSTRUCTIONYEAR > YEAR 
                                          | is.na(CADCONSTRUCTIONYEAR), 
                                          CONSTRUCTIONYEAR, CADCONSTRUCTIONYEAR)) |>  
  dplyr::mutate(-IMPUTED_YEAR)
```
\normalsize
Remove duplicate records, we only take as repeated instance an asset located in the same place with the same price
\tiny
```{r Remove duplicates, message=F, warning=F}
assets.sf <- assets.sf |>
  dplyr::group_by(PERIOD,LATITUDE, LONGITUDE, PRICE, .drop=T) |>
  dplyr::filter(row_number() == 1) |>
  dplyr::ungroup()
```
\normalsize

We turn our regular dataframe to a spatial features data frame, we project the source coordinates.
\tiny
```{r Create spatial features, message=F, warning=F}
#
# Createsf and reproject
#

assets.sf <- assets.sf |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = projcrs_src) |>
  st_transform(projcrs_dest)

#
# Save coordinates in a separate dataframe for later use
#

assets.coordinates = as_tibble(st_coordinates(assets.sf)) |>
  dplyr::mutate(IDX = row_number())

#
# Remove geometry (turn sf into ordinary a dataframe)
#
st_geometry(assets.sf) <-  NULL
```
\normalsize

# Price obfuscation

We will modify slighlty the prices, as price are almost-continuous in a two-step process:

* Randomize price with +/- 2.5%
* Adversisers usually round up figures to multiples of 10 (rent) or 1000 euros (sale), we align prices to multiples of 1000 (sales) and 10 (rent)
\tiny
```{r Transform prices, message=F, warning=F, echo=F}
#
# changes randomly the price with this amount
#
price_random_pct = 2.5

#
# Set the rounding grain
#
round_grain = list('sale' =  1000.0, 'rent'= 10)

#
# Perform the price adjustment
#
assets.sf <- assets.sf |>
  dplyr::mutate(VAR = runif(nrow(.),
                            -price_random_pct/100, price_random_pct/100.0),
                #
                # Add random disturbance and realign
                #
                PRICE = round(((1+VAR) * PRICE)/round_grain[[OPERATION]])*round_grain[[OPERATION]],
                PRICE = round(PRICE),
                NEW_UNITPRICE = PRICE / CONSTRUCTEDAREA) |>
  select(-VAR)
```
\normalsize

# Relocation of coordinates

We move the source coordinates with a radius between 30 and 60 meters:

* Take a random angle 0 - 360 degrees
* Scale the unitary radius to a value between 30 and 60 meters

The process makes sure listings are not moved our their neighborhood boundary (for this matter we use the zoning schema used on www.idealista.com). 
\tiny
```{r Set coordinates and locations for the relocation process, message=F, warning=F}
#
# Set distance parameters
#

MIN = 30
MAX = 60
DIFF = MAX - MIN

#
# Set location codes for each record: join between polygon areas and listings' cooordinates
#

coordinates.sf <- assets.coordinates |>
    st_as_sf(coords = c("X", "Y"), crs = projcrs_dest) |>
    st_join(polygons.sf, left = T)

#
# Remove spatial features from the dataframe
#

st_geometry(coordinates.sf) <- NULL

#
# It should not necessary if polygons do not overlap, just in case they do we take the first location 
#  the listing is located within
#

coordinates.sf <- coordinates.sf |>
  dplyr::group_by(IDX, .drop=T) |>
  dplyr::summarize(LOCATIONID = min(LOCATIONID), groups = NULL) |>
  dplyr::arrange(IDX)

#
# We assign the idealista's location code (called LOCATIONID)
#

assets.coordinates$LOCATIONID = coordinates.sf$LOCATIONID
```
\normalsize

We move the coordinates, bearing in mind we must keep the item in the same polygon that it was, if not the case try get new coordinates again (to a maximum of 10 fold, in practice it needed 2 iteration maximum). 
\tiny
```{r Regenerate coordinates, message=F, warning=F}
#
# Set a maximum reprocessing iteration
# a reprocess takes care of all items that are moved out from their original idealista polygons (LOCATIONID)
#
MAX_ITERATIONS = 10

#
# Init the process
#
pending.coordinates = assets.coordinates
new.coordinates = NULL
iteration = 1

cat(paste0('Start process objective ', nrow(pending.coordinates), ' rows\n'))

while(nrow(pending.coordinates) > 0 | iteration > MAX_ITERATIONS){

    cat(paste0('\tPerforming relocation iteration ', iteration, 
             ' pending rows: ',nrow(pending.coordinates),'\n' ))

  #
  # Relocate coordinates has two steps:
  #
  #    1) Get a random angle alpha
  #    2) Calculate a distance between d in [MIN, MAX]
  #    3) Set new coordinates at (X + cos(alpha) * d, Y + sin(alpha) * d) 
  #
  candidate = pending.coordinates |>
      dplyr::mutate(ANGLE_RADIANS = runif(nrow(.), 0, 360) * pi / 180) |>
      dplyr::mutate(DX = cos(ANGLE_RADIANS) * (MIN + DIFF * runif(nrow(.), 0, 1)),
                    DY = sin(ANGLE_RADIANS) * (MIN + DIFF * runif(nrow(.), 0, 1))) |>
      dplyr::mutate(NEW_X = round(X + DX),
                    NEW_Y = round(Y + DY)) |>
      dplyr::select(-ANGLE_RADIANS, -DX, -DY)

  #
  # Append the new location to the data frame`
  #
  candidates.sf = candidate |>
      st_as_sf(coords = c("NEW_X", "NEW_Y"), crs = projcrs_dest) |>
      st_join(polygons.sf) |>
      dplyr::rename(LOCATIONID = LOCATIONID.x,
                    NEW_LOCATIONID = LOCATIONID.y)

  # Take locationid use the min
  candidates.sf <- candidates.sf |>
    dplyr::group_by(IDX, LOCATIONID, .drop = T) |>
    dplyr::summarize(NEW_LOCATIONID = min(NEW_LOCATIONID), .groups = NULL) |>
    dplyr::arrange(IDX)

  candidate$NEW_LOCATIONID = candidates.sf$LOCATIONID

  # Add created end redo the pending
  pending.coordinates <- candidate |>
    dplyr::filter(NEW_LOCATIONID != LOCATIONID & !is.na(LOCATIONID))

  done.coordinates  <- candidate |>
    dplyr::filter((NEW_LOCATIONID == LOCATIONID) | is.na(LOCATIONID))

  if (is.null(new.coordinates)){
    new.coordinates = done.coordinates
  } else {
    new.coordinates <- new.coordinates |>
      dplyr::bind_rows(done.coordinates)
  }
  cat(paste0('\tPerformed relocation iteration ', iteration, 
             ' total rows: ',nrow(done.coordinates),'\n-------\n' ))
  iteration = iteration + 1

}
new.coordinates <- new.coordinates |>
  mutate(MOVED_DISTANCE = sqrt((NEW_X - X)^2 + (NEW_Y - Y)^2))
rm(pending.coordinates, done.coordinates, candidates.sf, candidate)
```
\normalsize

Assign the new coordinates to the listings.
\tiny
```{r Reproject coordinates, message=F, warning=F, echo=F}
assets.sf <- assets.sf |>
  dplyr::mutate(LONGITUDE = new.coordinates$NEW_X,
                LATITUDE = new.coordinates$NEW_Y) |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = projcrs_dest) |>
  st_transform(projcrs_src) 
```
\normalsize
\tiny
```{r Assign final new coordinates}
#
# Save coordinates
#
assets.coordinates = as_tibble(st_coordinates(assets.sf))

#
# Remove geometry
#
st_geometry(assets.sf) <-  NULL

#
# Reasign coordinates once geom was removed
#
assets.sf <- assets.sf |>
  dplyr::mutate(LONGITUDE = assets.coordinates$X,
                LATITUDE = assets.coordinates$Y)

```
\normalsize

# Save results

Create final dataframe and save results
\tiny
```{r Save results, message=F, warning=F}
#
# Save results
#
assets.sf |>
  fwrite(file = paste0(output_path, iconv(file_name,from="UTF-8",to="ASCII//TRANSLIT")), 
         sep = ";",
         compress = "gzip")
```
\normalsize